---
title: "KTheobald_Livesession5Assignment"
author: "K. Theobald"
date: "2/10/2019"
output:
  html_document:
    keep_md: true
---
## R Markdown

## Github Directory : https://github.com/TheobaldK/MSDS-6306-Homework.git

##**Question 1**##

### Backstory: Your client is expecting a baby soon.  However, he is not sure what to name the child.  Being out of the loop, he hires you to help him figure out popular names.  He provides for you raw data in order to help you make a decision.

### 1.	Data Munging: Utilize yob2016.txt for this question. This file is a series of popular children’s names born in the year 2016 in the United States.  It consists of three columns with a first name, a gender, and the amount of children given that name.  However, the data is raw and will need cleaning to make it tidy and usable.

### a.	First, import the .txt file into R so you can process it.  Keep in mind this is not a CSV file.  You might have to open the file to see what you’re dealing with.  Assign the resulting data frame to an object, df, that consists of three columns with human-readable column names for each.


```{r, echo=TRUE, cache=TRUE}  
#set the working directory
setwd("~/Desktop/Kari/SMU /Doing Data Science/Homework")

#Import the .txt file into R. Assign the data frame as df with 3 column
babynames<- "yob2016.txt" 
df<-read.table(babynames, header=FALSE,  sep=";")
colnames(df)<-c(V1="Baby Names", V2="Gender", V3="Babies")
head(df)
```


### b.	Display the summary and structure of df
```{r, echo=TRUE, cache=TRUE}  
str(df)
```
### c.	Your client tells you that there is a problem with the raw file.  One name was entered twice and misspelled.  The client cannot remember which name it is; there are thousands he saw! But he did mention he accidentally put three y’s at the end of the name.  Write an R command to figure out which name it is and display it.

```{r, echo=TRUE, cache=TRUE}  
#Find the 3 "Y's" at the end of the name.
dup.name<-grep("yyy", df$`Baby Names`)
#show the duplicate name index
dup.name
#show the duplicate name
df$`Baby Names`[dup.name]
str(df)
```
### d.	Upon finding the misspelled name, please remove this particular observation, as the client says it’s redundant.  Save the remaining dataset as an object: y2016 

```{r, echo=TRUE, cache=TRUE}  
#Remove the duplicate name/observation with the extra y's and save the dataset as an object y2016.
y2016 <- df[-c(212),]
str(y2016)
```
##**Question 2**
### Data Merging: Utilize yob2015.txt for this question.  This file is similar to yob2016, but contains names, gender, and total children given that name for the year 2015.

### a.	Like 1a, please import the .txt file into R.  Look at the file before you do.  You might have to change some options to import it properly.  Again, please give the dataframe human-readable column names.  Assign the dataframe to y2015.  

```{r, echo=TRUE, cache=TRUE}  
#set working directory
setwd("~/Desktop/Kari/SMU /Doing Data Science/Homework")

#Import the .txt file into R. Assign the data frame as df with 3 column
babynames2015<- "yob2015.txt" 
y2015<-read.table(babynames2015, header=FALSE,  sep=",")
colnames(y2015)<-c(V1="Baby Names", V2="Gender", V3="Babies")
head(y2015)
```

### b.	Display the last ten rows in the dataframe.  Describe something you find interesting about these 10 rows.
```{r, echo=TRUE, cache=TRUE}  
#Show the last 10 rows of the 2015 baby name dataframe. 
tail(y2015, 10)
```

#### I find the last few names to be male names each with exactly five times. The baby names are also alphabetical. 


### c.	Merge y2016 and y2015 by your Name column; assign it to final.  The client only cares about names that have data for both 2016 and 2015; there should be no NA values in either of your amount of children rows after merging.

```{r, echo= TRUE, cache=TRUE}
#Merge the 2016 and 2015 data frames by the baby names. Exclude the NA's. 
head(y2015)
Final<-merge(x=y2015,y=y2016, by="Baby Names", all=FALSE)
str(Final)
```
##**Question 3.	Data Summary: Utilize your data frame object final for this part.**

###  a.	Create a new column called “Total” in final that adds the amount of children in 2015 and 2016 together.  In those two years combined, how many people were given popular names?

```{r, echo= TRUE, cache=TRUE}
#Add the total number of babies from 2015 and 2016.
Final$Total<-rowSums(Final[,c("Babies.x", "Babies.y")])
#Sum the total number of popular baby names. 
sum(Final$Total)
#Show there is a column called Total for the number of babies in 2015 and 2016. 
str(Final)
```
### b.	Sort the data by Total.  What are the top 10 most popular names?


```{r, echo= TRUE, cache=TRUE}
library(tidyr)
library(dplyr)
#Create the dataframe from Final and call it MostPopular. Show the most 10 popular baby names regardless of Gender. The highest number of babies with these names will be at the top. 
MostPopular<-Final[order(Final$Total, decreasing=TRUE),]
head(MostPopular, 10)
```
#### The top 10 most popular names are Emma, Olivia, Noah, Liam, Sophia, Ava, Mason, William, Jacob, Isabella.

### c.	The client is expecting a girl!  Omit boys and give the top 10 most popular girl’s names.
```{r, echo= TRUE, cache=TRUE}
#Changed the Gender.x and Gender.y to character strings. 
Final$Gender.x<-as.character(Final$Gender.x)
Final$Gender.y<-as.character(Final$Gender.y)
# Created a data frame CleanFinal, from merging y2015 and y2016 Baby Names and Gender. 
CleanFinal<-merge(y2015, y2016, union("Baby Names", "Gender"))
# Created a Total Column in the dataframe CleanFinal summing the columnes of Babies.x (2015) and Babies.y (2016).
CleanFinal$Total<-rowSums(CleanFinal[,c("Babies.x", "Babies.y")])
# Sorted the data frame by order, using the largest to smallest value (decreasing), for the Total Babies. 
CleanFinal<-CleanFinal[order(CleanFinal$Total, decreasing=TRUE),]
# Sorted the data Frame by order, using the alphabetical order starting with F for Females. 
CleanFinal<-CleanFinal[order(CleanFinal$Gender, decreasing=FALSE),]
# Recreated the data frame CleanFinal by only including the 3 columns of Baby Names, Gender and Total
CleanFinal<-data.frame(CleanFinal$`Baby Names`, CleanFinal$Gender, CleanFinal$Total)
# Rename the column names to Names, Gender and Babies.
names(CleanFinal)<- c("Names", "Gender", "Babies")
# Show the top ten female baby names in the CleanFinal data frame. 
head(CleanFinal, 10)
```
### d.	Write these top 10 girl names and their Totals to a CSV file.  Leave out the other columns entirely.
```{r, echo= TRUE, cache=TRUE}
# Write onlly the Top 10 girl names. Started by creating a data frame with only Names and number of Babies with these names. 
TopGirlNames<-data_frame(CleanFinal$Names, CleanFinal$Babies)
# Renamed the column headers.
names(TopGirlNames)<- c("Names", "Total Babies")
# Display only the top 10 names for the data frame. 
head(TopGirlNames, 10)
# Write to a csv file the data frame with only the first 10 rows and all columns, call the file Girlnames.csv and include the row names and no na's if there were any. There are no na's. 
write.csv(TopGirlNames[1:10,], file = "Girlnames.csv", row.names=TRUE, na="")
```


##**Question 4**	
####Upload to GitHub (10 points): Push at minimum your RMarkdown for this homework assignment and a Codebook to one of your GitHub repositories (you might place this in a Homework repo like last week).  The Codebook should contain a short definition of each object you create, and if creating multiple files, which file it is contained in.  You are welcome and encouraged to add other files—just make sure you have a description and directions that are helpful for the grader.

### Provide the R Session Info:
```{r, echo= TRUE, cache=TRUE}
sessionInfo()
```
